\documentclass[12pt,fleqn]{article}
%\documentclass[12pt,a4paper]{article}
\usepackage{natbib}
\usepackage{lineno}
%\usepackage{lscape}
%\usepackage{rotating}
%\usepackage{rotcapt, rotate}
\usepackage{amsmath,epsfig,epsf,psfrag}
\usepackage{setspace}
\usepackage{ulem}
\usepackage{xcolor}
\usepackage[labelfont=bf,labelsep=period]{caption} %for making figure and table numbers bold
%\usepackage{a4wide,amsmath,epsfig,epsf,psfrag}


\def\be{{\ensuremath\mathbf{e}}}
\def\bx{{\ensuremath\mathbf{x}}}
\def\bX{{\ensuremath\mathbf{X}}}
\def\bthet{{\ensuremath\boldsymbol{\theta}}}
\newcommand{\VS}{V\&S}
\newcommand{\tr}{\mbox{tr}}
%\renewcommand{\refname}{\hspace{2.3in} \normalfont \normalsize LITERATURE CITED}
%this tells it to put 'Literature Cited' instead of 'References'
\bibpunct{(}{)}{,}{a}{}{;}
\oddsidemargin 0.0in
\evensidemargin 0.0in
\textwidth 6.5in
\headheight 0.0in
\topmargin 0.0in
\textheight=9.0in
%\renewcommand{\tablename}{\textbf{Table}}
%\renewcommand{\figurename}{\textbf{Figure}}
\renewcommand{\em}{\it}

\begin{document}

\begin{center} \bf {\large ESTIMATING MULTI-SPECIES ABUNDANCE USING THERMAL VIDEO AND DIGITAL PHOTOGRAPHY: ICE-ASSOCIATED SEALS IN THE EASTERN BERING SEA}

\vspace{0.7cm}
Paul B. Conn$^{1*}$, Jay M. Ver Hoef$^1$, Brett T. McClintock$^1$, Erin E. Moreland$^1$, Josh M. London$^1$, Michael F. Cameron$^1$, and Peter L. Boveng$^1$
\end{center}
\vspace{0.5cm}

\rm
\small

\it $^1$NOAA-NMFS, Alaska Fisheries Science Center, National Marine Mammal Laboratory, 7600 Sand Point Way NE, Seattle, WA, 98115, U.S.A.\\

\rm \begin{flushleft}

\raggedbottom

\begin{center}
Appendix S1: Gibbs sampling and full conditional distributions for estimating animal abundance
from a combination of thermal imagery and digital photography
\bigskip
\end{center}

We use a hybrid Metropolis-Gibbs sampler \citep[cf.][]{GelmanEtAl2004} to iteratively sample from full conditional distributions
of model parameters.  For each iteration of the Markov chain, we sample full conditionals as follows (note that notation is defined in Table 1 of the article):

\begin{enumerate}
  \item Sample regression parameters describing species-landscape covariate relationships using
  \begin{equation*}
  [\boldsymbol{\beta}_s|\hdots]={\rm Normal}(({\bf X_s}^\prime {\bf X_s})^{-1}{\bf X_s}^\prime \boldsymbol{\nu}_s,(\tau_{\nu s}+\tau_\beta)^{-1}({\bf X}_s^\prime {\bf X})_s^{-1}).
  \end{equation*}
  This formulation implies an ${\rm MVN}({\bf 0},(\tau_\beta X_s^\prime X_s)^{-1})$ prior distribution
  on regression parameters, where $\tau_\beta$ is a prior precision parameter (we set $\tau_\beta=0.01$ in subsequent applications). \\

  \item Sample $\nu_{js}$ for sampled cells using a random walk Metropolis-Hastings step.  The full conditional
  for this update is proportional to
  \begin{equation*}
  N({\bf X}_s \boldsymbol{\beta}_s + \boldsymbol{\eta},\tau_s^{-1})\times
    {\rm Poisson}(G_{js}^{\rm obs};\lambda_{js}).
  \end{equation*} 
  A tuning phase can be employed to adapt the kernel for random walk proposals to be in the optimal [0.30,0.40] range \citep{GelmanEtAl2004}.
  \\

  \item Sample $\nu_{js}$ for unsampled cells directly as
    $\boldsymbol{\nu}_{s} \sim N({\bf X}_s \boldsymbol{\beta}_s + \boldsymbol{\eta},\tau_s^{-1})$ \\

  \item Update $\boldsymbol{\alpha}_s$ using a Gibbs step (recall that under dimension reduction, spatial random effects, $\boldsymbol{\eta}_s$, can be expressed as $\boldsymbol{\eta}_s={\bf K}\boldsymbol{\alpha}_s$).  Specifically, sample $\boldsymbol{\alpha}_s \sim {\rm MVN}(\boldsymbol{\mu}_s,\boldsymbol{\Sigma}_s)$, where $\boldsymbol{\Sigma}_s^{-1}={\bf K}^\prime {\bf K}\tau_{\nu s}+\tau_{\eta s}{\bf K}^\prime {\bf Q} {\bf K}$ and $\boldsymbol{\mu}_s=\boldsymbol{\Sigma}_s \tau_{\nu s} {\bf K}^\prime (\boldsymbol{\nu}_s - {\bf X}_s \boldsymbol{\beta}_s) $ \\

  \item Update $\tau_{\eta s}$ using a Gibbs step, where
  \begin{equation*}
  \tau_{\eta s} \sim {\rm Gamma}(0.5m + a,0.5 \boldsymbol{\alpha}^\prime
  {\bf K}^\prime {\bf Q} {\bf K}\boldsymbol{\alpha}+b).
  \end{equation*}
  This formulation implies that $\tau_{\eta s}$ has an ${\rm Gamma}(a,b)$ prior distribution. We set $a=1$ and $b=0.01$ in worked examples, which gives ample support to likely values (e.g. those in (0,500]) while maintaining an approximate uniform distribution near the origin. \\


  \item If estimated, $\tau_{\nu s}$ can also be updated using a Gibbs step when assuming a conjugate
      ${\rm Gamma}(a,b)$ prior.  The full conditional in this case is ${\rm Gamma}(0.5L+a,0.5 \boldsymbol{\Delta}_s^\prime \boldsymbol{\Delta}_s +b)$, where $\boldsymbol{\Delta}_s=\boldsymbol{\nu}_s-{\bf X}_s \boldsymbol{\beta}_s-\boldsymbol{\eta}_s$ \\

  \item For hotspots with $I_{ij}=0$, simulate species from a categorical distribution with probabilities proportional to $\lambda_{js}$, $s \in [1,S]$.  Simulate individual covariates for these observations directly from $f(\boldsymbol{\theta}_s)$ \\

  \item For hotspots with $I_{ij}=1$, update true species identity $S_{ij}$ using a Metropolis-Hastings step.  Species proposals are generated by sampling from $\rm{Uniform}\{1,2,\hdots,S\}$.  The full conditional distribution has contributions from the Poisson abundance model, the covariate model, and the species identification model.  Specifically, the full conditional distribution for species identity being $S_{ij}=s$ is proportional to
      \begin{equation*}
        \lambda_{js} \pi^{[O_{ij}|s]} \prod_k f_k(Z_{ijk} ; \boldsymbol{\theta}_s)
      \end{equation*} \\

  \item For each individual covariate, update parameters of covariate distributions.  For instance, if species $s$ group sizes are assumed to follow a zero-truncated ${\rm Poisson}(\theta_s)$ distribution (as we assume in worked examples), selection of a conjugate ${\rm Gamma}(a,b)$ prior distribution allows the parameter $\theta_s$ to be updated via a Gibbs step where
       \begin{equation*}
       \theta_s \sim {\rm Gamma}\left( \sum_j \sum_i \left[ Z_{ij1} I_{ij} -I_{ij} \right] + a,\sum_j \sum_i I_{ij} + b \right).
       \end{equation*} \\
       In worked examples we set $a=b=0.1$.

  \item Update thinning parameters, $p_{js}$, using an independence chain.  Since there is little information to estimate these parameters from survey data, our strategy is to use posterior predictions from an auxiliary analysis as both a prior distribution and proposal distribution for ${\bf p}$ during abundance analysis (see e.g. \texttt{Example}).  We update the vector of thinning parameters using a Metropolis-Hastings step, with full conditional distribution proportional to \begin{equation*}
      \prod_j \prod_s {\rm Poisson}(G_{js}^{\rm obs};A_j R_j p_{js} \exp(\nu_{js})).
      \end{equation*} 
      As proposals are iid, there is no need for tuning.\\

  \item Update species misclassification parameters, $\boldsymbol{\pi}$ using an independence chain.  As with
      updates of ${\bf p}$, we use an informative prior on $\boldsymbol{\pi}$ as the proposal distribution within a Metropolis-Hastings step.  The full conditional is then simply
      \begin{equation*}
        \prod_j \prod_s \prod_i (1-I_{ij})+I_{ij} \pi^{[O_{ij}|S_{ij}]}.
      \end{equation*}
          As proposals are iid, there is no need for tuning.
\end{enumerate}

We then compute posterior predictive loss for each model $m$ fit to the data as $D_m = G_m + P_m$,
where $G_m$ is a measure of posterior loss and $P_m$ is a penalty for variance.
For each model, we simulate $n$ replicate datasets by drawing from the joint posterior distribution and
using the same set of distributional assumptions we used to estimate parameters.  We then compute
\begin{equation*}
   G_m = \sum_j \parallel {\bf O}_{j}^{\rm rep} - {\bf O}_j^{\rm obs} \parallel ^2,
\end{equation*}
where ${\bf O}_{j}$ gives a vector summarizing the total number of observations in cell $j$ that were
 of each observation type; the superscript denotes whether the data were simulated ('rep') or observed
 ('obs').

A penalty for variance is then computed as
\begin{equation*}
   P_m = \sum_j \sum_k {\rm Var}(O_{jk}^{\rm rep}),
\end{equation*}
where variance is taken over simulation replicates and $k$ indexes observation type.

\renewcommand{\refname}{Literature Cited}
\bibliographystyle{JEcol}

\bibliography{master_bib}

\end{flushleft}
\end{document}

